---
title: "ROhdsiWebApi"
output: github_document
---

ROhdsiWebApi


[![Build Status](https://travis-ci.org/OHDSI/ROhdsiWebApi.svg?branch=master)](https://travis-ci.org/OHDSI/ROhdsiWebApi)
[![codecov.io](https://codecov.io/github/OHDSI/ROhdsiWebApi/coverage.svg?branch=master)](https://codecov.io/github/OHDSI/ROhdsiWebApi?branch=master)

## Introduction
An R package with functions to interact with OHDSI WebAPI web services, identified by baseUrl. This package is useful to use R to make GET, POST, PUT, DELETE API calls using R-script. This is very useful for reproducible research, and also to run long running studies in batch mode. An example of a WebApi endpoint is "http://server.org:80/WebAPI".

The package supports the following functional categories in WebApi.

```{r eval=TRUE, echo = FALSE}
library(magrittr) 
output <- ROhdsiWebApi:::.getStandardCategories() %>% dplyr::select(categoryFirstUpper) %>% dplyr::rename(Category = categoryFirstUpper) %>% dplyr::arrange() %>% dplyr::mutate(features = paste0("Functions for interfacing with ", Category, " in WebApi"))
colnames(output) <- colnames(output) %>% SqlRender::camelCaseToTitleCase()
output
```

The package follows a generic design pattern: 
+ Functions that go from WebApi to R: will return a R-object (e.g. list, data frame). The function will handle the conversion of non R-objects like JSON objects into R-objects in the background. The returned objects may be inspected in R. We have ensured that all returned objects are represented in the proper data types e.g. dates in are type date, instead of string, use of integer where appropriate etc.
+ Functions that go from R into WebApi: will start with a R-object (e.g. list, data frame). If the WebApi requires a JSON object, the function will handle the conversion of R-object into JSON in the background. 

The package follows a consistent naming conventions - with the identifier for the functional category clearly in the name. There are no direct reference to 'Atlas', and instead 'WebApi' is used = because the package is designed to interface directly with WebApi.

## Technology
ROhdsiWebApi is an R package.

## System Requirements
Requires R.

## Installation
In R, use the following commands to download and install ROhdsiWebApi:

```{r eval=TRUE, echo = FALSE}
remotes::install_github("OHDSI/ROhdsiWebApi")
```

## User Documentation
+ Documentation can be found on the [package website](https://ohdsi.github.io/ROhdsiWebApi).

PDF versions of the documentation are also available:
- Vignette: 
+ [Working With Cohorts](https://raw.githubusercontent.com/OHDSI/ROhdsiWebApi/master/inst/doc/WorkingWithCohorts.pdf)
- Package manual: 
+ [ROhdsiWebApi.pdf](https://raw.githubusercontent.com/OHDSI/ROhdsiWebApi/master/extras/ROhdsiWebApi.pdf)

## Support
- Developer questions/comments/feedback: <a href="http://forums.ohdsi.org/c/developers">OHDSI Forum</a>
- We use the <a href="https://github.com/OHDSI/ROhdsiWebApi/issues">GitHub issue tracker</a> for all bugs/issues/enhancements 
 
## Contributing
Read [here](https://ohdsi.github.io/Hades/contribute.html) how you can contribute to this package.
 
## License
ROhdsiWebApi is licensed under Apache License 2.0

## Development
ROhdsiWebApi is being developed in R Studio.

