---
title: "Insert Cohort Definitions Into Package"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    number_sections: yes
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Using ROhdsiWebApi}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

ROhdsiWebApi is part of [HADES](https://ohdsi.github.io/Hades).
# Introduction
From Package [Readme](https://github.com/OHDSI/ROhdsiWebApi/blob/master/README.md)

> ROhdsiWebApi maybe used to insert cohort definition json and sql from your Atlas (webapi) instance 
into your study R package. Two functions in this package support `insertCohortDefinitionInPackage` 
and `insertCohortDefinitionSetInPackage` . Please see documentation for the functions. This vignette 
focusses on the `insertCohortDefinitionsetInPackage` function. It depends on a reference file 
called CohortsToCreate.

> We can easily create a cohorts to create files as follows:

```{r eval=FALSE, echo = TRUE}

library(magrittr)
# Set up
baseUrl <- Sys.getenv("BaseUrl")
# list of cohort ids
cohortIds <- c(18345,18346)

# get specifications for the cohortIds above
webApiCohorts <-
        ROhdsiWebApi::getCohortDefinitionsMetaData(baseUrl = baseUrl) %>%
        dplyr::filter(.data$id %in% cohortIds)

cohortsToCreate <- list()
for (i in (1:nrow(webApiCohorts))) {
  cohortId <- webApiCohorts$id[[i]]
  cohortDefinition <-
    ROhdsiWebApi::getCohortDefinition(cohortId = cohortId, 
                                      baseUrl = baseUrl)
  cohortsToCreate[[i]] <- tidyr::tibble(
    atlasId = webApiCohorts$id[[i]],
    name = webApiCohorts$id[[i]],
    cohortName = stringr::str_trim(stringr::str_squish(cohortDefinition$name))
  )
}

cohortsToCreate <- dplyr::bind_rows(cohortsToCreate)

readr::write_excel_csv(x = cohortsToCreate, na = "", 
                       file = "inst/settings/CohortsToCreate.csv", 
                       append = FALSE)

```

> Once this file has been created, then use insert function as follows

```{r eval=FALSE, echo = TRUE}
# Insert cohort definitions from ATLAS into package -----------------------
ROhdsiWebApi::insertCohortDefinitionSetInPackage(
  fileName = "inst/settings/CohortsToCreate.csv",
  baseUrl = baseUrl,
  insertTableSql = TRUE,
  insertCohortCreationR = TRUE,
  generateStats = TRUE,
  packageName = 'yourFavoritePackage'
)
```
